### 2024-10-07 doc/Rogers_20XX/bat_guano.Rmd

---
title: A template for the *arxiv* style
authors:
  - name: Timothy J. Rogers
    thanks: 
    department: Department of Bioinformatics
    affiliation: University of North Carolina
    location: Charlotte, NC 28223
    email: troger50@charlotte.edu
  - name: Richard A. White
    department: Department of Bioinformatics
    affiliation: University of North Carolina
    location: Charlotte, NC 28223
    email: rwhit101@charlotte.edu
abstract: |
  Enter the text of your abstract here.
keywords:
  - guano
  - microbiome
  - bat
  - virome
  - metagenomics
bibliography: "`r rbbt::bbt_write_bib('references.bib', overwrite = TRUE)`" 
#bibliography: references.bib
biblio-style: unsrt
output: rticles::arxiv_article
---

# Introduction

Here goes an introduction text

# Headings: first level
\label{sec:headings}

You can use directly LaTeX command or Markdown text.

LaTeX command can be used to reference other section. See Section \ref{sec:headings}.
However, you can also use **bookdown** extensions mechanism for this.

## Headings: second level

You can use equation in blocks

$$
\xi _{ij}(t)=P(x_{t}=i,x_{t+1}=j|y,v,w;\theta)= {\frac {\alpha _{i}(t)a^{w_t}_{ij}\beta _{j}(t+1)b^{v_{t+1}}_{j}(y_{t+1})}{\sum _{i=1}^{N} \sum _{j=1}^{N} \alpha _{i}(t)a^{w_t}_{ij}\beta _{j}(t+1)b^{v_{t+1}}_{j}(y_{t+1})}}
$$

But also inline i.e $z=x+y$

### Headings: third level

Another paragraph. [@deleon_etal18] and see bob.

# Examples of citations, figures, tables, references
\label{sec:others}

The documentation for \verb+natbib+ may be found at

You can use custom blocks with LaTeX support from **rmarkdown** to create environment.

::: {.center latex=true}
  <http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}>
:::

Of note is the command \verb+\citet+, which produces citations
appropriate for use in inline text.

You can insert LaTeX environment directly too.

\begin{verbatim}
   \citet{hasselmo} investigated\dots
\end{verbatim}

produces

\begin{quote}
  Hasselmo, et al.\ (1995) investigated\dots
\end{quote}

\begin{center}
  \url{https://www.ctan.org/pkg/booktabs}
\end{center}

## Figures

You can insert figure using LaTeX directly.

See Figure \ref{fig:fig1}. Here is how you add footnotes. [^Sample of the first footnote.]

\begin{figure}
  \centering
  \fbox{\rule[-.5cm]{4cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  \caption{Sample figure caption.}
  \label{fig:fig1}
\end{figure}

But you can also do that using R.

```{r fig2, fig.cap = "Another sample figure"}
plot(mtcars$mpg)
```

You can use **bookdown** to allow references for Tables and Figures.

## Tables

Below we can see how to use tables.

See awesome Table~\ref{tab:table} which is written directly in LaTeX in source Rmd file.

\begin{table}
 \caption{Sample table title}
  \centering
  \begin{tabular}{lll}
    \toprule
    \multicolumn{2}{c}{Part}                   \\
    \cmidrule(r){1-2}
    Name     & Description     & Size ($\mu$m) \\
    \midrule
    Dendrite & Input terminal  & $\sim$100     \\
    Axon     & Output terminal & $\sim$10      \\
    Soma     & Cell body       & up to $10^6$  \\
    \bottomrule
  \end{tabular}
  \label{tab:table}
\end{table}

You can also use R code for that.

```{r}
knitr::kable(head(mtcars), caption = "Head of mtcars table")
```

## Lists

- Item 1
- Item 2 
- Item 3